---
title: "Problem Set: Explore One Variable"
output: html_notebook
---

***

### Diamonds

```{r}
library(ggplot2)
data(diamonds)

summary(diamonds)
?diamonds
```


***

### Price Histogram
```{r}
ggplot(aes(x = price), data = diamonds) + geom_histogram(bins = 15, fill = '#099DD9', color = 'black') + scale_x_continuous()
```


***

### Price Histogram Summary
```{r}
summary(diamonds$price)
```

***

### Daimond Counts
How many diamonds cost less than $500?
How many diamonds cost less than $250?
How many diamonds cost $15,000 or more?
```{r}
nrow(subset(diamonds, diamonds$price < 500))
nrow(subset(diamonds, diamonds$price < 250))
nrow(subset(diamonds, diamonds$price >= 15000))

```

***

### Cheaper Diamonds
```{r}
ggplot(aes(x = price), data = diamonds) + geom_histogram(bins = 30, fill = '#099DD9', color = 'black') + scale_x_continuous(breaks = seq(300, 1000, 50), limits = c(300, 1000))
```


***

### Price by Cut Histograms
```{r}
ggplot(aes(x = price), data = diamonds) + geom_histogram(bins = 15, fill = '#099DD9', color = 'black') + scale_x_continuous(breaks = seq(300, 2000, 500), limits = c(300, 2000)) + facet_wrap(~cut)
```


***

### Price by Cut
Which cut has the highest priced diamond?
Which cut has the lowest priced diamond?
Which cut has the lowest median price?
```{r}
by(diamonds$price, diamonds$cut, summary)
```


***

### Scales and Multiple Histograms
```{r}
ggplot(aes(x = price), data = diamonds) + geom_histogram(bins = 15, fill = '#099DD9', color = 'black') + scale_x_continuous(breaks = seq(300, 2000, 500), limits = c(300, 2000)) + facet_wrap(~cut, scales = 'free_y')
```

***

### Price per Carat by Cut
```{r}
ggplot(aes(x = price/carat), data = diamonds) + geom_histogram(bins = 20, fill = '#099DD0', color = 'black') + scale_x_log10() + facet_wrap(~cut, scales = 'free_y')
```


***

### Price Box Plots
```{r}
qplot(x = cut, y = price, data = diamonds, geom = 'boxplot') + coord_cartesian(ylim = c(0, 10000))
```
```{r}
qplot(x = color, y = price, data = diamonds, geom = 'boxplot') + coord_cartesian(ylim = c(0, 10000))
```
```{r}
qplot(x = clarity, y = price, data = diamonds, geom = 'boxplot') + coord_cartesian(ylim = c(0, 10000))
```


***

### Interquartile Range - IQR
```{r}
by(diamonds$price, diamonds$color, summary)
IQR(subset(diamonds, color == 'D')$price)
IQR(subset(diamonds, color == 'J')$price)
```


***

### Price per Carat Box Plots by Color
```{r}
ggplot(aes(x = color, y = price/carat), data = diamonds) + geom_boxplot() + coord_cartesian(ylim = c(0, 7500))
```


***

### Carat Frequency Polygon
```{r}
ggplot(aes(x = carat), data = diamonds) + geom_freqpoly(binwidth = .1)
table(diamonds$carat)
```


***

### Data Wrangling with R
https://s3.amazonaws.com/udacity-hosted-downloads/ud651/DataWranglingWithR.pdf
tidyr -a package that reshapes the layout of your data

dplyr - a package that helps you transform tidy, tabular data

https://www.rstudio.com/resources/cheatsheets/
***

### Gapminder Data

***

### Exploring Your Friends' Birthdays
How many people share your birthday? Do you know them?

Which month contains the most number of birthdays?

How many birthdays are in each month?

Which day of the year has the most number of birthdays?

Do you have at least 365 friends that have birthdays on everyday of the year?

```{r}
birth <- read.csv('birthdaysExample.csv')
birth$dates <- strptime(birth$dates, format = "%m/%d/%y")
birth$month <- strftime(birth$dates, format = "%m")
birth$day <- strftime(birth$dates, format = "%d")
birth$year <- strftime(birth$dates, format = "%y")

nrow(subset(birth, day == 25 & month == 10))

ggplot(aes(x = month), data = birth) + geom_histogram(stat = "count")

table(birth$month)

ggplot(aes(x = day), data = birth) + geom_histogram(stat = "count")

ggplot(aes(x = day), data = birth) + geom_histogram(binwidth = 1, stat = "count") + facet_wrap(~month)
```
*************************************************************************************************************

# Exploring Two Variables


***

### price vs. x
```{r}
ggplot(aes(x = x, y = price), data = diamonds) + geom_point()
```

***

### Findings - price vs. x
Price increases exponentially with x. There are outliers with x = 0.
***

### Correlations
```{r}
with(diamonds, cor.test(price, x))

with(diamonds, cor.test(price, y))

with(diamonds, cor.test(price, z))
```

***

### price vs. depth
```{r}
ggplot(aes(x = depth, y = price), data = diamonds) + geom_point()
```

***

### Adjustments - price vs. depth
```{r}
ggplot(aes(x = depth, y = price), data = diamonds) + geom_point(alpha = 1/100) + scale_x_continuous(breaks = seq(58, 64, 2), limits = c(58, 64))
```

***

### Typical Depth Range
***

### Correlation - price and depth
```{r}
with(diamonds, cor.test(depth, price))
```

***

### price vs. carat
```{r}
ggplot(aes(x = carat, y = price), data = diamonds) + geom_point() + xlim(0, quantile(diamonds$carat, .99)) + ylim(0, quantile(diamonds$price, .99))
```

***

### price vs. volume
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(aes(x = volume, y = price), data = diamonds) + geom_point()
```
```{r}
library(plyr)
count(diamonds$volume == 0)
detach("package:plyr", unload=TRUE)
```

***

### Findingns - price vs. volume

***

### Correlations on Subsets
```{r}
with(subset(diamonds, diamonds$volume > 0 & diamonds$volume < 800), cor.test(price, volume))
```

***

### Adjustments - price vs. volume
```{r}
ggplot(aes(x = volume, y = price), data = subset(diamonds, diamonds$volume > 0 & diamonds$volume < 800)) + geom_point(alpha = 1/50) + geom_smooth()
```

***

### Mean Price by Clarity
```{r}
library(dplyr)
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n()) 
head(diamondsByClarity)
```

***

### Bar Charts of Mean Price
```{r}
diamondsByColor <- diamonds %>%
  group_by(color) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n()) 
head(diamondsByColor)

library(gridExtra)

p1 <- ggplot(aes(x = clarity, y = mean_price), data = diamondsByClarity) + geom_col()

p2 <- ggplot(aes(x = color, y = mean_price), data = diamondsByColor) + geom_col()

grid.arrange(p1, p2, ncol = 1)
```

***

### Trends in Mean Price
As clarity improves, mean price of the diamond decreases. SI2, although not the worst clarity, is the most expensive category. As diamond color improves, the price decreases.
***

### Gapminder Revisited

*************************************************************************************************************

# Exploring Many Variables

***

### Price Histograms with Facet and Color
```{r}
ggplot(aes(x = price), data = diamonds) + facet_grid( ~ color) + geom_histogram(aes(color = cut)) + scale_x_log10()
```

***

### Price vs. Table Colored by Cut
```{r}
ggplot(aes(x = table, y = price), data = diamonds) + geom_point(aes(color = cut)) + scale_color_brewer(type = 'qual') + scale_x_continuous(breaks = seq(50, 80, 1), limits = c(50, 80))
```

***

### Price vs. Volume and Diamond Clarity
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z

ggplot(aes(x = volume, y = price), data = subset(diamonds, volume <= quantile(volume, .99))) + geom_point(aes(color = clarity)) + scale_color_brewer(type = 'div') + scale_y_log10()
```

***

### Proportion of Friendships Initiated
```{r}
pf$prop_initiated <- pf$friendships_initiated / pf$friend_count

pf$prop_initiated[is.na(pf$prop_initiated)] <- 0
```

***

### prop_initiated vs. tenure
```{r}
pf$year_joined <- floor(2014 - pf$tenure / 365)
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))

ggplot(aes(x = tenure, y = prop_initiated), data = subset(pf, pf$tenure > 0 )) + geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Smoothing prop_initiated vs tenure
```{r}
ggplot(aes(x = tenure, y = prop_initiated), data = subset(pf, pf$tenure > 0 )) + geom_smooth(aes(color = year_joined.bucket))
```

***

### Largest Group Mean prop_initiated
```{r}
tapply(pf$prop_initiated, pf$year_joined.bucket, summary)
```

***

### Price/Carat Binned, Faceted, and Colored
```{r}
ggplot(aes(x = cut, y = price / carat), data = diamonds) + geom_jitter(aes(color = color)) + scale_color_brewer(type = 'div') + facet_wrap( ~ clarity, ncol = 3)
```

